Liveness-Based Garbage Collection for Lazy Languages

### 2016-04-17 paper/paper.tex
First revision with Sanyal sir

In  this work  we show  that  static analysis  of heap  data can  help
garbage  collectors   in  reclaiming   more  garbage.    The  specific
contributions of this paper are:
\added{
\begin{itemize}
\item {\color{Myblue}We propose a liveness based GC scheme for a lazy
  first order functional language and prove its correctness.
  Thomas~\cite{Thomas19951} describes a copying garbage collector for
  Three Instruction Machine (TIM) that only traces live variables in a
  closure's environment, i.e., the variables that are useful for
  evaluating the closure.  Their notion of liveness of variables is
  close to the classical notion of liveness, and does not take into
  account liveness of parts of structures (say, only the spine of a
  list).  To the best of our knowledge, it is the only earlier work on
  GC that takes liveness for closures into account.  All previous
  attempts~\cite{shaham01heap, ran.shaham-sas03, shaham02estimating,
    asati14lgc, karkare06effectiveness} have involved either
  imperative or eager functional languages.  The novelty of our
  approach is in handling of liveness of closures, an issue that
  arises due to laziness at runtime.}

@@

\item The  analysis results  in a set  of context-free  grammars along
  with  a fixed  set  of non-context-free  productions.  The  decision
  whether to copy a cell during  GC translates to a membership problem
  for such  grammars.  Earlier research assumed  the undecidability of
  this membership question and  used an over-approximation to overcome
  it.  In this paper, we provide  a formal proof of the undecidability
  of this problem.

@@

>>>>

In  this work  we show  that  static analysis  of heap  data can  help
garbage  collectors   in  reclaiming   more  garbage.    The  specific
contributions of this paper are:
\added{
\begin{itemize}
\item {\color{Myblue}We  propose an interprocedural liveness  based GC
  scheme  for a  lazy first  order functional  language and  prove its
  correctness.  To the  best of our knowledge, this is  the first work
  that uses  the results  of an  interprocedural liveness  analysis to
  garbage    collect    both     evaluated    data    and    closures.
  Thomas~\cite{Thomas19951} describes a  copying garbage collector for
  the  Three  Instruction  Machine  (TIM)  that  only  preserves  live
  closures in a function's environment (also called a frame). However,
  in the absence of  details, it is not clear whether  a) the scope of
  the method is interprocedural, and b) it handles algebraic datatypes
  like lists (the original design of TIM did not).  All other previous
  attempts~\cite{shaham01heap,  ran.shaham-sas03,  shaham02estimating,
    asati14lgc, karkare06effectiveness} involved  either imperative or
  eager functional languages.}
%% The novelty of our approach is in handling of
%%   liveness  of closures,  an  issue  that arises  due  to laziness  at
%%   runtime.}

@@

\item The  analysis results  in a set  of context-free  grammars along
  with  a fixed  set  of non-context-free  productions.  The  decision
  whether to copy a cell during  GC translates to a membership problem
  for such grammars.  Earlier, this decision problem was assumed to be
  undecidable, and the context-free grammar was over approximated by a
  regular  grammar.   In  this  paper,  we  provide  a  proof  of  the
  undecidability of this problem.

@@

We  now return  back  to the  problem of  computing  liveness using
\Lfonly\  and a  concrete  demand $\sigma$.  Although  the process  is
explained through an example, it can be generalized easily.
